Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2025-12-12 response_round1.tex

%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
\usepackage{booktabs} %for the table in the analysis section
\usepackage{enumerate} %for the short algorithm
\usepackage{subfigure} %for making plots with multiple images
\usepackage{tabularx} %for the spacing in tabular
\usepackage{newtxmath} %for making indicator function
\usepackage{changepage} %for adjusting the margins
\usepackage{dsfont} %for getting the indicator function
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\1}[1]{\mathds{1}\left[#1\right]}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{amsmath}

\usepackage{xcolor}
\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
\newcommand{\spencer}[1]{{\color{red} Response: #1}}

\begin{document}
\Huge
\noindent Response to reviewers
\\~\\
\normalsize
We are grateful to the three reviewers and their insightful comments and 
questions. We apologize for the omission of responses to some comments in our
resubmission on 12/5/25, and are embarassed for having somehow overlooked that.
We thank the editor for the chance to correct that mistake, and we have 
corrected it.
\\~\\
Overall, the reviewer comments
made a huge impact on improving the content of
the manuscript. In the most recent revision, we are particularly grateful for 
the questions on uncertainty propagation. Obviously the explanation we gave
for how uncertainty was propagated wasn't clear enough, so in this revision
we reworked sections 3.7 and 3.8 to clarify the procedure.
Specifics are found below and in the 
revised manuscript. Reviewer comments are \emph{italicized}, our responses to 
each reviewer's comments are in normal script, 
and specific changes/additions to the main manuscript are 
indented.
\\~\\

@@

\Huge
\noindent AE1
\normalsize
\\~\\

% \emph{
% I think the authors did a great job in this newly submitted version and 
% addressed most of the earlier raised comments in a nice and convincing way. 
% The paper is well written and interesting to read. 
% }
% I only have two remaining comments.

\begin{enumerate}[-]

  \item \emph{Section 3.7. End of first paragraph: ``..., but propagation of 
  uncertainty   in ILI forecasts was still permitted in the estimating of 
  hospitalisation forecasts.``
  }
  \emph{
  Please clarify more explicitly how uncertainty was propagated between the two 
  model components. My understanding is that this was achieved by incorporating 
  not only point estimates of the parameters but also samples from the predictive 
  distribution of the ILI model. If this is correct, please state it clearly 
  already at this stage and consider adding a brief discussion of the advantages 
  and disadvantages of this approach.}
  \\~\\
  To clarify how uncertainty was propagated, we consolidated sections
  3.7 and 3.8, reordered the main ideas, and removed redundant information. We
  also added some advantages/disadvantages of our approach. The changes now
  read:
  \begin{quote}
        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
        \label{sec:implementation_posterior}

        Rather than combine the ILI and hospitalization models into a single model,
        forecasts were made via a two-stage fitting process. In the first stage,
        the ILI model (1) and hospitalization model
        (6) were fit separately via Bayesian posterior updating. 
        In the second stage, the two model fits were
        fused together to produce hospitalization 
        forecasts. Future ILI 
        forecasts were obtained from the ILI forecast model in (\ref{eq:ili_post}).
        Here 
        $p(\xi | \textbf{ILI}_w)$ 
        is the posterior distribution of the ILI model (1)
        and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$.
        The ILI forecast model was obtained by integrating over the 
        parameters 
        $\xi = (\pi$, $\kappa$, $\sigma^2_{\gamma}$, $\sigma^2_{\gamma_W})$.
        The desired forecasts are for weeks $w + i$ 
        where $i$ is a positive integer, and $\widetilde{ILI}_{w + i}$ represents the 
        predicted or forecasted ILI at $i$ weeks in the future.
        Note that
        in this section
        the parameter subscript for season $s$ is dropped to simplify notation.

@@

        % \begin{equation}
        %     \label{eq:ili_post}
        %     p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
        %     p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
        %     \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
        %     \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
        %     d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
        % \end{equation}

        \begin{equation}
        \tag{7}
            \label{eq:ili_post}
            p(\widetilde{ILI}_{w +i} | \textbf{ILI}_w) = \int  
            p(\widetilde{ILI}_{w + i} | \xi) 
            p(\xi | \textbf{ILI}_w) 
            d\xi
        \end{equation}
        % \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
        % \spencer{Isn't that captured in $d \pi_s$?}
        Also in the first stage, the posterior distribution of the hospitalization 
        model (6) given observed
        hospitalization data $\textbf{H}_w$ and $\textbf{ILI}_w$ was obtained.
        The posterior distribution is denoted as
        $p(\psi | \textbf{H}_w, \textbf{ILI}_w)$.
        In the second stage,
        the ILI forecast model in (\ref{eq:ili_post}) was then combined with 
        the hospitalization posterior distribution
        to produce a hospitalization forecast $\widetilde{H}_{w + i}$ 
        for week $w + i$.

        Both the ILI forecast model in (\ref{eq:ili_post}) and the hospitalization
        posterior distributions were estimated via Markov chain Monte Carlo 
        (MCMC) sampling.
        From the fit ILI model, $K$ MCMC draws
        $(\widetilde{ILI}_{w +i}^{(1)}, ..., \widetilde{ILI}_{w +i}^{(K)})$
        from the forecast model were produced, 
        $K$ draws $(\psi^{(1)}, ..., \psi^{(K)})$
        from the hospitalization model posterior distribution
        were produced. Each of the
        ILI forecast and hospitalization draws was then plugged into 
        the hospitalization model
        (6) to produce the approximate hospitalization forecast
        distribution in the bottom line of (\ref{eq:hosp_post_pred}). 
        This has 
        similarities to the procedure used in section 3.2.2 of 
        \cite[]{sahu2014hierarchical} who use it to propagate the uncertainty of 
        temperature forecasts into forecast model for hospitalizations.
        Here $k \in \{1, ..., K\}$, 
        and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
        In the analyses in sections 4 and 5
        we set $K = 50,000$ draws.

@@

        \begin{equation}
        \tag{8}
        \begin{aligned}
        \label{eq:hosp_post_pred}
          &p(\widetilde{H}_{w + i} | \textbf{H}_w, \textbf{ILI}_w)\\
          &= \int p(\widetilde{H}_{w +i} | 
               \psi, \widetilde{ILI}_{w +i}, \textbf{H}_w)
               p(\psi | \textbf{H}_w, \textbf{ILI}_w) d \psi d\widetilde{ILI} \\
          &\approx \frac{1}{K} p(\widetilde{H}_{w + i}^{(k)} | \psi^{(k)},  
                              \tilde{ILI}_{w +i}^{(k)}, \textbf{H}_w, \textbf{ILI}_w) 
        \end{aligned}
        \end{equation}
        % \[
        %  \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
        %                                   \alpha_{1s}^{(k)} 
        %                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
        %                                   \alpha_{2s}^{(k)} 
        %                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
        %                                   \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
        %                                   \epsilon_{s,w + i}^{(k)}
        % \]
        % where $k \in \{1, ..., K\}$. 
        % The hospitalization draws were then combined into
        % $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
        % from which
        % the probabilistic hospitalization forecast was estimated.
        % In practice we set
        % $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.

        % \[ \begin{array}{rl}
        % p(\tilde{H}_w | H_w, ILI_w) 
        % &=               \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | H_w, ILI_w)  d\tilde{ILI}_w \\
        % &\approx  \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | ILI_w)  d\tilde{ILI}_w
        % \end{array} \]

        In the two-stage scheme, the posterior distribution of the ILI
        model (1) parameters was not directly influenced by the 
        hospitalization data and the posterior distribution of the hospitalization model 
        (6) was not directly influenced by ILI model 
        parameters. Hence there was limited feedback between the
        two components, but the propagation of uncertainty in ILI forecasts was still 
        accomplished in estimating hospitalization forecasts through
        the forecast model (\ref{eq:hosp_post_pred}).
        This two-stage modeling is reminiscent of the Bayesian cut 
        described in \cite[]{plummer2015cuts} and \cite[]{nott2023bayesian}. Cutting a model
        into two component models or ``modules'' may be justified for several reasons
        including to reduce time to fit the model, avoiding mixing issues of posterior
        sampling distributions, improving predictive performance or to prevent
        model misspecification in one model to influence the other 
        \cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
        lee2017rigorous, plummer2015cuts}. A drawback
        from cutting a model is potential loss of useful information. 
        In our current 
        application for example, the ILI model parameters receive no potentially 
        useful information from hospitalization data.

        The decision to cut the forecast model into two component models
        was made for three major reasons.
        When fitting the fully Bayesian joint model, the time to fit a single model 
        using
        MCMC sampling and
        obtain forecasts was sometimes many hours whereas fitting the two models
        separately took a fraction of the time. When assessing convergence of model
        parameters, it was rare that posterior sampling chains for the joint model
        showed convergence even after 55,000 iterations, whereas the component model 
        fits gave no signs which raised concerns about convergence. And most importantly,
        the forecasts from the component models generally outperformed those of the 
        joint model in terms of minimizing proper scoring rules.
        \\~\\
  \end{quote}
  \emph{
  \item I recommend including the statement raised by the reviewer in the first 
  submission—that WIS approximates CRPS (ideally supported with a reference)—and 
  therefore leads to essentially the same conclusions. I think the motivation for 
  using WIS instead of CRPS is well explained, especially given that WIS is the 
  primary evaluation measure used in the FluSight competition} 
  \\~\\

  The following statements were added on page 17.
  \begin{quote}

@@

    We also removed comparisons with any competing models with less than 
    80\% submissions and further clarified the need to use relative scores.
    The following 2 snippets were added on page 20.

    \begin{quote}

      During the FluSight competition, most participating teams would occasionally 
      miss a forecast submission for some week, location, or horizon.
      The FluSight 
      models which are used for comparison are the
      13 non-ensemble models which submitted forecasts for at least 80\% of 
      FluSight targets during the 2023 season. 
      \\~\\

      Because forecast submissions for some targets were occasionally missed
      by participating teams, for any two models there was often a mismatch between
      targets for which forecasts were included or excluded. This makes a
      simple average of scores an imperfect measure for comparison.
    \end{quote}

@@

\newpage

\Huge
\noindent R2
\normalsize
\\~\\
% Forecasting Influenza Hospitaliztions Using a Bayesian Hierarchical Nonlinear 
% Model with Discrepancy
% 
% 
\emph{This manuscript seems to be a very reasonable attempt at forecasting influenza
induced hospitalizations. The
manuscript also puts the methodology in the context of the FluSight competition
organised by the CDC in the US.
Apologies, I have found it hard to review as it is conceptually very hard to do
so in my opinion.}

\emph{Firstly,
I do not find comfortable modelling the proportion, rather a natural outcome to
model will be the number of hospitalization using a Generalized Linear model
(Poisson distribution).}

The influenza like illness (ILI) data model in equation (1) is indeed 
proportion data, and is reported as the proportion of patients visiting
a provider who display flu like symptoms. This is stated in the introduction
on page 2, but in the revision we restate it again as below on page 4

\begin{quote}
An ILI case is defined as a 
``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
and/or a sore throat,'' and ILI data is reported as the proportion 
or percentage of patients who display these symptoms.
\end{quote}

For that reason, we argue that 
model (1) is appropriate for ILI data. 
To clarify that the purpose of this paper is
to use historically accurate ILI forecasts as a predictor of hospitalizations,
we added the following two statements; the first in the introduction on page
3 and the second in section 3 on page 7.

\begin{quote}
The contribution of this manuscript is to introduce a two component 
Bayesian modeling framework 
for modeling HHS hospitalization forecasts where hospitalization data and 
years of ILI data are used to inform forecast models. 
As is shown in this manuscript,
ILI is a good predictor of hospitalizations, 
and years of experience have led
to effective ILI forecast methods. Thus producing accurate ILI forecasts 
and using them to inform hospitalizations greatly assists in producing
accurate hospitalization forecasts.
\end{quote}

\begin{quote}
Presented in this section is methodology for forecasting hospitalizations
using ILI forecasts as a predictor of hospitalization forecasts. 
The availability of nearly a decade of ILI data along with accurate forecast
models allows us to make accurate forecasts of ILI which are used as a linear
predictor for hospitalization forecasts.
\end{quote}

Hospitalization data which we are trying to forecast is
count data. Thus a Poisson is certainly a natural modeling choice. To 
keep the model simple, however, we modeled hospitalizations using a normal
distribution, a location-scale t distribution,
or log transformed hospitalizations and modeled the transformed data with
a normal linear model. See model (6).
This choice was made because the forecasts were accurate,
the modeling is simpler, and fitting the model using a Poisson was much slower
making it more difficult to meet the forecast deadlines. This argument was 
added on page 12.

\begin{quote}
Another natural choice would have been to
model hospitalizations with a Poisson response generalized linear model. 
However, a generalized linear model adds additional complexity and we found that
fitting the model required significantly more time than
model (6) making it more difficult to produce forecasts
for all 54 locations and meet the FluSight 
submission deadline which was roughly 12 hours after data were released.
Also to avoid complexity, the authors of \cite[]{sahu2014hierarchical} opted to 
forecast hospitalizations using a normal response model rather than a Poisson
response model.
\end{quote}

\emph{Secondly, I find it un-intuitive to model the
flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model
we focus on two aspects: the systematic part and the random part. Here, as far
as  I understand the systematic part is the ASG function. The random
part has several components, which are hidden in Equation (5). My difficulty in
modelling using the ASG
function lies in my inability to relate any meaningful meteorological variables
(which are very much drivers of the seasonal flu outbreaks). See for example
the paper,  ``A hierarchical Bayesian model for improving short-term forecasting
of hospital demand by including meteorological information`` by  SK Sahu,
B Baffour, PR Harper, JH Minty, C Sarran
Journal of the Royal Statistical Society, Series A (2014).}

The ASG function, and likewise the susceptible-infectious-recovered (SIR)
mechanistic model, are used to model
ILI data because of known annual trajectory, the relationship to 
mathematical infection disease models, and most importantly
the success they've had in 
producing accurate ILI forecasts.
ILI data and forecasts are used as 
linear predictors for hospitalizations in model (6).
Neither model (1) nor (6) preclude the use of additional linear predictors
for prediciting ILI or hospitalizations, but if additional predictors 
were to be used it would make more sense to add them to model (6).
We added the following two statements to discuss this important point.
The first is near the beginning of section 3.5 on page 12, and the second is 
in the discussion on page 24.

\begin{quote}
In model (6), the data is limited to the ILI and 
hospitalization data provided by FluSight, but including additional
linear predictors would be straightforward. In fact it has been shown that 
meteorological predictors and forecasts such as temperature can lead
to improved forecasts as well as geographic variables such as patient
age and sex \cite[]{sahu2014hierarchical}.  
\end{quote}

\begin{quote}
Additionally, it has been shown that meteorological variables
such as local temperature forecasts and geographic variables such as age can 
lead to improved hospitalization forecasts \cite[]{sahu2014hierarchical}. 
A straightforward and useful extension of the methodology herein would be to 
collect such variables and incorporate them as added hospitalization 
predictors.
\end{quote}

% In spite of the above criticisms, 
\emph{I appreciate the hierarchical modeling 
approach of fitting the two models separately in components. But I wonder how 
the uncertainty has been propagated from one component model to another. 
Cutting the feedback idea is well understood but one still needs to integrate 
uncertainty to get an overall measure of the uncertainty in the forecasts. 
I did not clearly understand this issue from my reading of  the paper. I can 
suggest one way to propagate this uncertainty is by using MCMC samples 
(and methods), see for example, the paper,  ``A rigorous statistical framework 
for spatio-temporal pollution prediction and estimation of its long-term impact 
on health`` by 
D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).}
\\~\\
We appreciated this comment and the included reference. We used the reference
to assist in our new explanation, and made sure to cite it. 
To clarify how uncertainty was propagated, we consolidated sections
  3.7 and 3.8, reordered the main ideas, and removed redundant information. We
  also added some advantages/disadvantages of our approach. The changes now
  read:
  \begin{quote}
        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
        \label{sec:implementation_posterior}

### 2025-12-12 manuscript_res.tex
adjust manuscript

To better inform public decision making regarding the flu epidemic, in 2013 the 
United States Centers for Disease Control and Prevention (CDC) organized a 
national flu forecasting competition, also known as FluSight 
\cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. 
The competition lasted from the fall of 2013 through spring of 2014, and
originally over a dozen teams of researchers from academic and industry 
backgrounds participated in FluSight by contributing their own forecast models. 
Besides the 2020-21 season, FluSight has been operated annually and 
researchers outside 
the CDC have been invited to participate. From hereon, we denote a 
flu season by the year in which the outbreak began, i.e. the 2020 season 
season outbreak began in 2020 and lasted until spring 2021.
Initially the target data for 
forecasts was influenza-like illness (ILI) data. ILI is the proportion of 
patients who meet a healthcare provider and who display flu like symptoms, 
and ILI data has been available at the state and national level since the 2010 
flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
The collaborative ILI forecasting effort has led to a number of modeling 
developments in flu forecasting 
\cite[see references therein for more examples]{mcandrew2021adaptively, 
osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their 
introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most 
commonly used flu forecasting models into four classes including mechanistic 
models based on differential equation compartmental models, agent based models 
based on population simulation, machine learning/regression models including 
data driven machine learning and statistical models, and data assimilation 
models which are constructed by assimilating mechanistic models into a 
probabilistic framework. An additional forecast modeling class used in FluSight 
involves the combination of several forecasts into a single ensemble forecast, 
which has been shown to perform well relative to individual models 
\cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.

@@

The contribution of this manuscript is to introduce a two component 
Bayesian modeling framework 
for modeling HHS hospitalization forecasts where hospitalization data and 
years of ILI data are used to inform forecast models. The first modeling 
component is a hierarchical model of ILI data. 
The second is a model of hospitalization 
data with ILI as a predictive covariate.
Herein we use Bayesian ILI models similar to those in Osthus et al. 
\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
The model of Osthus et al. \cite[]{osthus2019dynamic} is a combined data 
assimilation and statistical regression model which involves a compartmental 
model in a probabilistic framework. Their model also includes an additional 
component for capturing a systematic discrepancy between the deterministic 
part of the model and the actual data, an idea which was first introduced by 
Kennedy and O'Hagan \cite[]{kennedy2001bayesian}. The model in Ulloa 
\cite[]{ulloa2019} is a Bayesian hierarchical regression model with an 
underlying function intended to capture the trajectory of the seasonal 
ILI data. Herein, we provide a framework under which discrepancy modeling may 
be used along with a general function modeling ILI data. Modeling discrepancy
introduced challenges which required thoughtful selection of model 
constraints and prior distributions to be used effectively.

@@

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network collects 
information on respiratory illness from outpatient visits to health care 
providers. Each week, over 3,400 outpatient health care providers in all 50 
US states, PR, 
DC, and the US Virgin Islands report the total number of outpatient 
visits along with the number of ILI cases. An ILI case is defined as a 
``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
and/or a sore throat.'' Prior to the 2021 season, the definition included 
``without a known cause other than influenza'' \cite[]{cdcfluview2023}. Because 
other illnesses such as COVID-19, RSV, and the common cold may induce similar 
respiratory symptoms, ILI may include patients infected with some disease 
other than influenza. To know whether or not a sick patient is infected with 
influenza requires a laboratory test.

@@

Weekly hospital admission data, which was used as the object of FluSight 
forecasting for the 
2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
Patient Impact and Hospital Capacity by State Timeseries}. 
In February 2022 it became mandatory for 
all hospitals to report the number of COVID-19 and influenza hospitalizations, 
and since then reporting of hospitalizations has become widespread. These data 
were updated every Wednesday and Friday according to NHSN guidelines 
\cite[]{healthdata2024covidts}. Weekly influenza hospitalization data is
defined as the number of newly hospitalized patients with a confirmed diagnosis 
of influenza. From hereon, influenza hospitalizations
will be referrred to as hospitalizations.

The bottom of figure \ref{fig:us_ili} shows the weekly national 
hospitalizations for the 2022 and 2023 flu seasons. These plots show a 
trend similar to the ILI plots in that during the early 
weeks of the season hospitalizations are low, but they increase in the fall 
to a peak and then decrease until the flu outbreak ends. For both the 
2022 and 2023 seasons, the hospitalizations peaked during the same week as 
ILI, and in 2023 that peak occurred during the holiday week 22. Additional 
plots of ILI data and hospitalization
data for all 50 US states are included in the supplementary material 
\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories 
similar to those in figure \ref{fig:us_ili}.

@@

% \jarad{Why not start with the ILI model from Osthus et al? It makes a lot more sense to me to start with what has already been done and then expand upon what they have done. Later you have a section called ILI Model, but isn't this the ILI model?}

The typical behavior of the ILI data to begin low, rise to a peak, and then 
fall motivates the use of a nonlinear function which follows 
a similar trajectory. Compartmental models are standard mathematical models 
used for modelling disease outbreaks which may capture ILI behavior. 
One important compartmental model is the 
susceptible-infectious-recovered (SIR) 
disease transmission model, which has been used by some to model 
ILI data \cite[]{osthus2019dynamic, allen2017primer}.  
The SIR is a mechanistic model used to capture transmission of a particular
disease, and since ILI may be influenced by more 
diseases than influenza, the disease transmission described by the SIR model
can only partially contribute to the true ILI data-generating mechanism
\cite[]{osthus2019dynamic}.

@@

The second component for forecast modeling is based on the relationship between 
hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
This is an example of an autoregressive model with exogenous variables where 
the autoregressive lag is one (ARX(1)) 
\cite[]{raftery2010online,ljung1987system}.

@@

In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for 
week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed 
according to some distribution $D_s$ with mean parameter 0, scale parameter 
$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ which 
may be used as degrees of freedom
when $D_s$ belongs to the location-scale t (LST) family.

@@

Additional prior constraints were made to improve parameter identifiability. 
Just as was done in Osthus et al. \cite[]{osthus2019dynamic},
we set the initial value of the 
susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
priors. In a previous version of this manuscript, when fitting an ASG model
with discrepancy components,
rather than estimating all parameters through posterior updating,
identifiability was encouraged by setting
the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
to be equal to its MLE. In this updated manuscript, 
however, we use empirical Bayes and center the first component of the 
hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
all seasons and set the variance parameter so the prior was tight around the
mean. This led to improved mixing of posterior draws and better forecasts.
The 
specific prior distributions assigned to each parameter are listed in the 
supplementary materials.

@@

% \cite[]{wadsworth2024bas}.

>>>>

To better inform public decision making regarding the flu epidemic, in 2013 the 
United States Centers for Disease Control and Prevention (CDC) organized a 
national flu forecasting competition, also known as FluSight 
\cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. 
The competition lasted from the fall of 2013 through spring of 2014, and
originally over a dozen teams of researchers from academic and industry 
backgrounds participated in FluSight by contributing their own forecast models. 
Besides the 2020-21 season, FluSight has been operated annually and 
researchers outside 
the CDC have been invited to participate. From hereon, we denote a 
flu season by the year in which the outbreak began, i.e. the 2020 
season outbreak began in 2020 and lasted until spring 2021.
Initially the target data for 
forecasts was influenza-like illness (ILI) data. ILI is the proportion of 
patients who meet a healthcare provider and who display flu like symptoms, 
and ILI data has been available at the state and national level since the 2010 
flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
The collaborative ILI forecasting effort has led to a number of modeling 
developments in flu forecasting 
\cite[see references therein for more examples]{mcandrew2021adaptively, 
osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their 
introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most 
commonly used flu forecasting models into four classes including mechanistic 
models based on differential equation compartmental models, agent based models 
based on population simulation, machine learning/regression models including 
data driven machine learning and statistical models, and data assimilation 
models which are constructed by assimilating mechanistic models into a 
probabilistic framework. An additional forecast modeling class used in FluSight 
involves the combination of several forecasts into a single ensemble forecast, 
which has been shown to perform well relative to individual models 
\cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.

@@

The contribution of this manuscript is to introduce a two component 
Bayesian modeling framework 
for modeling HHS hospitalization forecasts where hospitalization data and 
years of ILI data are used to inform forecast models. 
As is shown in this manuscript,
ILI is a good predictor of hospitalizations, 
and years of experience have led
to effective ILI forecast methods. Thus producing accurate ILI forecasts 
and using them to inform hospitalizations greatly assists in producing
accurate hospitalization forecasts.
The first modeling 
component is a hierarchical model of ILI data. 
The second is a model of hospitalization 
data with ILI as a predictive covariate.
Herein we use Bayesian ILI models similar to those in Osthus et al. 
\cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
The model of Osthus et al. \cite[]{osthus2019dynamic} is a combined data 
assimilation and statistical regression model which involves a compartmental 
model in a probabilistic framework. Their model also includes an additional 
component for capturing a systematic discrepancy between the deterministic 
part of the model and the actual data, an idea which was first introduced by 
Kennedy and O'Hagan \cite[]{kennedy2001bayesian}. The model in Ulloa 
\cite[]{ulloa2019} is a Bayesian hierarchical regression model with an 
underlying function intended to capture the trajectory of the seasonal 
ILI data. Herein, we provide a framework under which discrepancy modeling may 
be used along with a general function modeling ILI data. Modeling discrepancy
introduced challenges which required thoughtful selection of model 
constraints and prior distributions to be used effectively.

@@

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network collects 
information on respiratory illness from outpatient visits to health care 
providers. Each week, over 3,400 outpatient health care providers in all 50 
US states, PR, 
DC, and the US Virgin Islands report the total number of outpatient 
visits along with the number of ILI cases. An ILI case is defined as a 
``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
and/or a sore throat,'' and ILI data is reported as the proportion 
or percentage of patients who display these symptoms.
Prior to the 2021 season, the definition included 
``without a known cause other than influenza'' \cite[]{cdcfluview2023}. Because 
other illnesses such as COVID-19, RSV, and the common cold may induce similar 
respiratory symptoms, ILI may include patients infected with some disease 
other than influenza. To know whether or not a sick patient is infected with 
influenza requires a laboratory test.

@@

Weekly hospital admission data, which was used as the object of FluSight 
forecasting for the 
2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
Patient Impact and Hospital Capacity by State Timeseries}. 
In February 2022 it became mandatory for 
all hospitals to report the number of COVID-19 and influenza hospitalizations, 
and since then reporting of hospitalizations has become widespread. These data 
were updated every Wednesday and Friday according to NHSN guidelines 
\cite[]{healthdata2024covidts}. Weekly influenza hospitalization data is
defined as the number of newly hospitalized patients with a confirmed diagnosis 
of influenza. From hereon, influenza hospitalizations
will be referred to as hospitalizations.

The bottom of figure \ref{fig:us_ili} shows the weekly national 
hospitalizations for the 2022 and 2023 flu seasons. These plots show a 
trend similar to the ILI plots in that during the early 
weeks of the season hospitalizations are low, but they increase in the fall 
to a peak and then decrease until the flu outbreak ends. For both the 
2022 and 2023 seasons, the hospitalizations peaked during the same week as 
ILI, and in 2023 that peak occurred during the holiday week 22. Additional 
plots of ILI data and hospitalization
data for all 50 US states are included in the supplementary material 
\cite[]{wadsworth2025bas}. These plots show flu outbreak trajectories 
similar to those in figure \ref{fig:us_ili}.

@@

% \jarad{Why not start with the ILI model from Osthus et al? It makes a lot more sense to me to start with what has already been done and then expand upon what they have done. Later you have a section called ILI Model, but isn't this the ILI model?}
Presented in this section is methodology for forecasting hospitalizations
using ILI forecasts as a predictor of hospitalization forecasts. 
The availability of nearly a decade of ILI data along with accurate forecast
models allows us to make accurate forecasts of ILI which are used as a linear
predictor for hospitalization forecasts.

The typical behavior of the ILI data to begin low, rise to a peak, and then 
fall motivates the use of a nonlinear function which follows 
a similar trajectory. Compartmental models are standard mathematical models 
used for modeling disease outbreaks which may capture ILI behavior. 
One important compartmental model is the 
susceptible-infectious-recovered (SIR) 
disease transmission model, which has been used by some to model 
ILI data \cite[]{osthus2019dynamic, allen2017primer}.  
The SIR is a mechanistic model used to capture transmission of a particular
disease, and since ILI may be influenced by more 
diseases than influenza, the disease transmission described by the SIR model
can only partially contribute to the true ILI data-generating mechanism
\cite[]{osthus2019dynamic}.

@@

The second component for forecast modeling is based on the relationship between 
hospitalizations and ILI and is defined in (\ref{eq:final_hosp}).
This is an example of an autoregressive model with exogenous variables where 
the autoregressive lag is one (ARX(1)) 
\cite[]{raftery2010online,ljung1987system}. 
In model (\ref{eq:final_hosp}), the data is limited to the ILI and 
hospitalization data provided by FluSight, but including additional
linear predictors would be straightforward. In fact it has been shown that 
meteorological predictors and forecasts such as temperature can lead
to improved forecasts as well as geographic variables such as patient
age and sex \cite[]{sahu2014hierarchical}.

@@

In model (\ref{eq:final_hosp}), $H_{s,w}$ is the number of hospitalizations for 
week $w$ in season $s$, $\epsilon_{s,w}$ is an error term distributed 
according to some distribution $D_s$ with mean parameter 0, scale parameter 
$\sigma_{\epsilon_s}$, and the additional parameter $\omega_s$ which 
may be used as degrees of freedom
when $D_s$ belongs to the location-scale t (LST) family. 
In the simulation study performed in section \ref{sec:simulation2} 
$D_s$ belongs to 
the normal distribution family, and for for analyzing the 2023 hospitalization
data presented in section \ref{sec:analysis} models were considered where 
$D_s$ was LST or where hospitalization counts were log transformed 
and the transformed
data were modeled as normally distributed.
Another natural choice would have been to
model hospitalizations with a Poisson response generalized linear model. 
However, a generalized linear model adds additional complexity and we found that
fitting the model required significantly more time than
model (\ref{eq:final_hosp}) making it more difficult to produce forecasts
for all 54 locations and meet the FluSight 
submission deadline which was roughly 12 hours after data were released.
Also to avoid complexity, the authors of \cite[]{sahu2014hierarchical} opted to 
forecast hospitalizations using a normal response model rather than a Poisson
response model.

@@

Additional prior constraints were made to improve parameter identifiability. 
Just as was done in Osthus et al. \cite[]{osthus2019dynamic},
we set the initial value of the 
susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
priors. In a previous version of this manuscript, when fitting an ASG model
with discrepancy components,
rather than estimating all parameters through posterior updating,
identifiability was encouraged by setting
the parameter $\lambda_s$ from (\ref{eq:asg_function_rep})
to be equal to its MLE. In this updated manuscript, 
however, we use empirical Bayes and center the first component of the 
hierarchical prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
all seasons and set the variance parameter so the prior was tight around the
mean. This led to improved mixing of posterior draws and better forecasts.
The 
specific prior distributions assigned to each parameter are listed in the 
supplementary materials.

@@

% \cite[]{wadsworth2025bas}.
